Windowed least-squares model reduction for dynamical systems

### 2019-10-08 section3/general_form.tex
proofing

\section{Windowed least-squares framework}\label{sec:tclspg} 
This section
outlines the proposed \methodNameLower\ 
(\methodAcronym) framework. In contrast to (1) the Galerkin approach, 
which minimizes the (time-continuous) FOM ODE residual at a time instance, 
(2) the LSPG approach, 
which minimizes
the (time-discrete) FOM O$\Delta$E residual 
over a time step, and (3) the ST-LSPG approach,  
which minimizes
the  FOM O$\Delta$E residual 
over the entire time interval, the proposed \methodAcronym\ method sequentially minimizes the 
time-continuous FOM ODE residual over a series of arbitrarily defined
\textit{time windows}. The formulation is compatible with both \spatialAcronym\ and \spaceTimeAcronym\ trial subspaces. 
In this section, we start by outlining the \methodAcronym\ formulation  
for a general space--time trial subspace. We then examine the \spatialAcronym\ 
trial subspace, followed by the \spaceTimeAcronym\ trial subspace. In each case, we derive the stationary conditions 
associated with the associated residual-minimization problems.
 %The proposed method can be formulated equivalently from
%two viewpoints. The first formulation corresponds to a sequence of
%minimization problems for the generalized coordinates that minimize the FOM
%ODE residual over each time slab. We refer to this as the 
%\textit{residual-minimization viewpoint}. The second viewpoint corresponds to
%formulating an optimal control problem for the Galerkin ROM. In this
%viewpoint, the \methodAcronym\ method is formulated as an optimal control
%problem in where the objective is to find a ``controller" that minimizes the
%residual of the Galerkin ROM. We refer to this viewpoint as the
%\textit{optimal control viewpoint.}  This section outlines the derivation of the \methodAcronym\ method
%from both the residual minimization and optimal control viewpoints.

%\subsection{Formulation of the Time-Continuous LSPG Method} In contrast to
%the LSPG approach, which minimizes the fully discrete residual over a
%time-step, the present work proposes minimizing the time-continuous FOM ODE
%residual over time slabs. The proposed method can be formulated equivalently
%from two viewpoints. The first approach formulates the problem from a
%standard calculus of variations viewpoint and is analogous to the classic
%residual minimization process used in the LSPG method (but here performed at
%the continuous level). The second viewpoint corresponds to formulating an
%optimal control problem in Lagrange form (which is a specific instance of an
%optimal control problem of Bolza type). In this viewpoint, the
%\methodAcronym\ is formulated as an optimal control problem in where the
%objective is to find a ``controller" that minimizes the residual of the
%Galerkin ROM. We now outline both of these (equivalent) formulations. 
%
\subsection{Windowed least-squares for general space--time trial subspaces} 
We begin by introducing a (potentially nonuniform) partition of the time domain $[0,T]$
into $\nslabs$ non-overlapping windows $[\timeStartArg{n} ,
\timeEndArg{n}]\subseteq[0,T]$ of length $\DeltaSlabArg{n}\defeq\timeEndArg{n} -
\timeStartArg{n}$, $n=1,\ldots,\nslabs$ such that 
$\timeStartArg{1} = 0$, $\timeEndArg{\nslabs} = T$, and
$\timeStartArg{n+1} = \timeEndArg{n}$,
$n=1,\ldots,\nslabs-1$; Fig.~\ref{fig:slab_fig} depicts this partitioning.
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid.pdf} 
\caption{Partitioning of the time
domain into time windows.} 
\label{fig:slab_fig} 
\end{centering} 
\end{figure}
Over the $n$th time window, we approximate the FOM ODE solution as 
$\approxstate^n(t)\approx \stateFOM(t)$,
$t\in[\timeStartArg{n},\timeEndArg{n}]$, which is enforced to reside in 
the $n$th space--time trial subspace such that
\begin{equation}
\approxstate^n \in \stspace^n \subseteq \RR{N} \otimes \timeSpaceArg{n} , \qquad  n = 1,\ldots,\nslabs,
\end{equation}
where $\stspaceArg{n}$ is the space--time trial subspace over the $n$th window, $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on
$[\timeStartArg{n},\timeEndArg{n}]$ (i.e., $\timeSpaceArg{n} = \{f\,|\,f:[\timeStartArg{n},\timeEndArg{n}]\rightarrow\RR{}\}$)
and
$\approxstate^n \in\RR{\fomdim}\otimes\timeSpaceArg{n}$. For notational purposes, we additionally define the (spatial) trial subspaces at the start of each window as
$$\approxstateArg{n}{\timeStartArg{n}} \in \stspaceBoundStartArg{n} \subset \stspaceArg{n}, \qquad n=1,\ldots,\nslabs. 
%, \qquad 
%\approxstateArg{n}{\timeEndArg{n}} \in \stspaceBoundEndArg{n} \subset \stspaceArg{n}.
$$ 
To outline \methodAcronym, we now define the objective functional over the $n$th window:
\begin{equation}\label{eq:obj}
\begin{split} \mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArgt{n}{t} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
%&\vcentcolon \RR{\fomdim} \times [\timeStartArg{n},\timeEndArg{n}] \rightarrow
%\RR{}, \end{align}
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n}  \rightarrow
\RR{}_+, 
\end{split}
\end{equation}
%\begin{align}\label{eq:obj} \mathcal{J}^n &\vcentcolon \state \mapsto
%\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \resid(\state)^T
%\stweightingMat \resid(\state) d\tau \\ &\vcentcolon \RR{\fomdim} \times
%[0,T] \rightarrow \RR{}.  \end{align}
where $\stweightingMatArg{n} \equiv \stweightingMatOneTArg{n}
\stweightingMatOneArg{n} \in \RR{\fomdim \times \fomdim}$ denotes a
symmetric positive semi-definite matrix that can enable hyper-reduction, for
example. 
%For notational simplicity, we employ the same weighting matrix for all time windows.

>>>>

\section{Windowed least-squares approach}\label{sec:tclspg} 
This section
outlines the proposed \methodNameLower\ 
(\methodAcronym) \approachKwd. In contrast to (1) the Galerkin approach, 
which minimizes the (time-continuous) FOM ODE residual at a time instance, 
(2) the LSPG approach, 
which minimizes
the (time-discrete) FOM O$\Delta$E residual 
over a time step, and (3) the ST-LSPG approach,  
which minimizes
the  FOM O$\Delta$E residual 
over the entire time interval, the proposed \methodAcronym\ method sequentially minimizes the 
time-continuous FOM ODE residual over a series of arbitrarily defined
\textit{time windows}. The formulation is compatible with both \spatialAcronym\ and \spaceTimeAcronym\ trial subspaces. 
In this section, we start by outlining the \methodAcronym\ formulation  
for a general space--time trial subspace. We then examine the \spatialAcronym\ 
trial subspace, followed by the \spaceTimeAcronym\ trial subspace. In each case, we derive the stationary conditions 
associated with the associated residual-minimization problems.
 %The proposed method can be formulated equivalently from
%two viewpoints. The first formulation corresponds to a sequence of
%minimization problems for the generalized coordinates that minimize the FOM
%ODE residual over each time slab. We refer to this as the 
%\textit{residual-minimization viewpoint}. The second viewpoint corresponds to
%formulating an optimal control problem for the Galerkin ROM. In this
%viewpoint, the \methodAcronym\ method is formulated as an optimal control
%problem in where the objective is to find a ``controller" that minimizes the
%residual of the Galerkin ROM. We refer to this viewpoint as the
%\textit{optimal control viewpoint.}  This section outlines the derivation of the \methodAcronym\ method
%from both the residual minimization and optimal control viewpoints.

%\subsection{Formulation of the Time-Continuous LSPG Method} In contrast to
%the LSPG approach, which minimizes the fully discrete residual over a
%time-step, the present work proposes minimizing the time-continuous FOM ODE
%residual over time slabs. The proposed method can be formulated equivalently
%from two viewpoints. The first approach formulates the problem from a
%standard calculus of variations viewpoint and is analogous to the classic
%residual minimization process used in the LSPG method (but here performed at
%the continuous level). The second viewpoint corresponds to formulating an
%optimal control problem in Lagrange form (which is a specific instance of an
%optimal control problem of Bolza type). In this viewpoint, the
%\methodAcronym\ is formulated as an optimal control problem in where the
%objective is to find a ``controller" that minimizes the residual of the
%Galerkin ROM. We now outline both of these (equivalent) formulations. 
%
\subsection{Windowed least-squares for general space--time trial subspaces} 
We begin by introducing a (potentially nonuniform) partition of the time domain $[0,T]$
into $\nslabs$ non-overlapping windows $[\timeStartArg{n} ,
\timeEndArg{n}]\subseteq[0,T]$ of length $\DeltaSlabArg{n}\defeq\timeEndArg{n} -
\timeStartArg{n}$, $n=1,\ldots,\nslabs$ such that 
$\timeStartArg{1} = 0$, $\timeEndArg{\nslabs} = T$, and
$\timeStartArg{n+1} = \timeEndArg{n}$,
$n=1,\ldots,\nslabs-1$; Fig.~\ref{fig:slab_fig} depicts this partitioning.
\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.0cm 5cm 0cm 3cm},clip,width=1.0\textwidth]{figs/time_grid.pdf} 
\caption{Partitioning of the time
domain into time windows.} 
\label{fig:slab_fig} 
\end{centering} 
\end{figure}
Over the $n$th time window, we approximate the FOM ODE solution as 
$\approxstate^n(t)\approx \stateFOM(t)$,
$t\in[\timeStartArg{n},\timeEndArg{n}]$, which is enforced to reside in 
the $n$th space--time trial subspace such that
\begin{equation}
\approxstate^n \in \stspace^n \subseteq \RR{N} \otimes \timeSpaceArg{n} , \qquad  n = 1,\ldots,\nslabs,
\end{equation}
where $\stspaceArg{n}$ is the space--time trial subspace over the $n$th window and $\timeSpaceArg{n}$ denotes the set of real-valued functions acting on
$[\timeStartArg{n},\timeEndArg{n}]$ (i.e., $\timeSpaceArg{n} = \{f\,|\,f:[\timeStartArg{n},\timeEndArg{n}]\rightarrow\RR{}\}$)
%and
%$\approxstate^n \in\RR{\fomdim}\otimes\timeSpaceArg{n}$. 
For notational purposes, we additionally define the (spatial) trial subspaces at the start of each window as
$$\approxstateArg{n}{\timeStartArg{n}} \in \stspaceBoundStartArg{n} \subset \stspaceArg{n}, \qquad n=1,\ldots,\nslabs. 
%, \qquad 
%\approxstateArg{n}{\timeEndArg{n}} \in \stspaceBoundEndArg{n} \subset \stspaceArg{n}.
$$ 
To outline \methodAcronym, we now define the objective functional over the $n$th window:
\begin{equation}\label{eq:obj}
\begin{split} \mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArgt{n}{t} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
%&\vcentcolon \RR{\fomdim} \times [\timeStartArg{n},\timeEndArg{n}] \rightarrow
%\RR{}, \end{align}
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n}  \rightarrow
\RRplus, 
\end{split}
\end{equation}
%\begin{align}\label{eq:obj} \mathcal{J}^n &\vcentcolon \state \mapsto
%\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \resid(\state)^T
%\stweightingMat \resid(\state) d\tau \\ &\vcentcolon \RR{\fomdim} \times
%[0,T] \rightarrow \RR{}.  \end{align}
where $\stweightingMatArg{n} \equiv \stweightingMatOneTArg{n}
\stweightingMatOneArg{n} \in \RR{\fomdim \times \fomdim}$ denotes a
symmetric positive semi-definite matrix that can enable hyper-reduction, for
example. 
%For notational simplicity, we employ the same weighting matrix for all time windows.

### 2019-10-08 section3/space_time.tex
proofing

%We now briefly consider \spaceTimeAcronym\ trial subspaces. 
The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
the FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$
with $\approxstateArgnt{n} \in \stspaceSTArg{n}$, where
\begin{equation}\label{eq:st_sttrialspace}
 \stspaceSTArg{n} \defeq
 \Range{\stbasisArgnt{n}} + \stateInterceptSTArg{n}\otimes \onesFunctionArg{n} .
\end{equation}
Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{0} = \bz$ is the space--time trial basis matrix \textit{function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation. 
For \spaceTimeAcronym\ trial subspaces, we take the affine transformation to be $\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for $n=2,\ldots,\nslabs$ and $\stateInterceptSTArg{1} = \stateFOMIC$; this definition of the trial subspace automatically satisfies the boundary conditions.
At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as,
 \begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},
\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window. 
\begin{comment}
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptSTArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
\KTC{Remove all commas before equations unless it makes sense gramatically.
Add the $n$ superscript to $x_{ref}$}
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
%where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

@@

%Inserting the approximation~\eqref{eq:stapprox} into the optimization problem~\eqref{eq:tclsrm}, the
%constrained minimization problem over the $n$th time slab can be re-written as,
%\begin{equation}\label{eq:obj_gen_slab_spacetime} \stgenstate^n =
%\underset{\stgenstatey}{\text{argmin }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \end{equation}
%%\begin{equation}\label{eq:obj_gen_slab} \genstate^n(t) =
%%\underset{\genstatey}{\text{argmin }} \mathcal{J}^n
%%\bigg(\decoder\big(\genstatey(\tau)\big) \bigg) , \end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC
% & n=1. \end{cases} \end{equation}
\subsubsection{Stationary conditions} 
The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial subspaces is as follows: in \spaceTimeAcronym\ trial subspaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial subspaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdim} \otimes \timeSpaceArg{n}$. 
Thus, in the \spaceTimeAcronym\ case the optimization problem is no longer minimizing a functional with respect to a function, but now is minimizing a function with 
respect to vector.  As such, the first-order optimality conditions can be
	derived using standard calculus. Differentiating the objective with respect
	to $\stgenstatey$ and setting equal to zero yields 
\begin{equation}\label{eq:st_stationary}
 \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
\stateInterceptSTArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},t) \bigg) dt = \bz.\end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC &
%n=1. \end{cases} \end{equation}
For \spaceTimeAcronym\ trial subspaces, the stationary conditions comprise an algebraic system of integral equations, as opposed to a system of differential equations.

>>>>

%We now briefly consider \spaceTimeAcronym\ trial subspaces. 
The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
the FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$
with $\approxstateArgnt{n} \in \stspaceSTArg{n}$, where
\begin{equation}\label{eq:st_sttrialspace}
 \stspaceSTArg{n} \defeq
 \Range{\stbasisArgnt{n}} + \stateInterceptSTArg{n}\otimes \onesFunctionArg{n} \subseteq \RR{\fomdim} \otimes \timeSpaceArg{n}.
\end{equation}
Here $\stbasisArg{n}{t} \in \RR{\fomdim \times \stdimArg{n}}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{\timeStartArg{n}} = \bz$ is the space--time trial basis matrix \textit{function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$ provides the affine transformation. 
For \spaceTimeAcronym\ trial subspaces, we take the affine transformation to be $\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for $n=2,\ldots,\nslabs$ and $\stateInterceptSTArg{1} = \stateFOMIC$; this definition of the trial subspace automatically satisfies the boundary conditions.
At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as,
 \begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},
\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window. 
\begin{comment}
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptSTArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
\KTC{Remove all commas before equations unless it makes sense gramatically.
Add the $n$ superscript to $x_{ref}$}
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
%where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

@@

%Inserting the approximation~\eqref{eq:stapprox} into the optimization problem~\eqref{eq:tclsrm}, the
%constrained minimization problem over the $n$th time slab can be re-written as,
%\begin{equation}\label{eq:obj_gen_slab_spacetime} \stgenstate^n =
%\underset{\stgenstatey}{\text{argmin }} \mathcal{J}^n \bigg( \stbasisArgnt{n} \stgenstatey + \stateInterceptArg{n}  \bigg) , \end{equation}
%%\begin{equation}\label{eq:obj_gen_slab} \genstate^n(t) =
%%\underset{\genstatey}{\text{argmin }} \mathcal{J}^n
%%\bigg(\decoder\big(\genstatey(\tau)\big) \bigg) , \end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC
% & n=1. \end{cases} \end{equation}
\subsubsection{Stationary conditions} 
The key difference between \spaceTimeAcronym\ and \spatialAcronym\ trial subspaces is as follows: in \spaceTimeAcronym\ trial subspaces the generalized coordinates comprise a vector in $\RR{\stdimArg{n}}$, while in the \spatialAcronym\ trial subspaces the generalized coordinates comprise a \textit{vector of functions}, i.e., the generalized coordinates live in $\RR{\romdimArg{n}} \otimes \timeSpaceArg{n}$. 
Thus, in the \spaceTimeAcronym\ case the optimization problem is no longer minimizing a functional with respect to a function, but now is minimizing a function with 
respect to vector.  As such, the first-order optimality conditions can be
	derived using standard calculus. Differentiating the objective with respect
	to $\stgenstatey$ and setting equal to zero yields 
\begin{equation}\label{eq:st_stationary}
 \intSlabArg{n} \bigg[ \stbasisDotArg{n}{t}^T  - \stbasisArg{n}{t}^T \bigg[\frac{\partial
\velocity}{\partial \stateyDiscrete} (\stbasisDotArg{n}{t} \stgenstateArg{n} +                    
\stateInterceptSTArg{n},t)\bigg]^T  \bigg] \stweightingMatArg{n} \bigg( \stbasisDotArg{n}{t} \stgenstateArg{n}  - \velocity (\stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},t) \bigg) dt = \bz.\end{equation}
%subject to the boundary conditions 
%\begin{equation}\label{eq:bcs_spacetime}
%\stbasisArg{n}{\timeStartArg{n}}\stgenstateArg{n} = \begin{cases} 
%\mathbb{P}^n\stbasisArg{n-1}{\timeEndArg{n-1}}\stgenstateArg{n-1}  & n = 2,\ldots,\nslabs, \\ 
%\mathbb{P}^n \stateFOMIC &
%n=1. \end{cases} \end{equation}
For \spaceTimeAcronym\ trial subspaces, the stationary conditions comprise an algebraic system of equations, as opposed to a system of differential equations.

### 2019-10-08 section3/tclsrm_formulation.tex
proofing

\subsection{\methodAcronym\ summary}
This section outlined the \methodAcronym\ method for model reduction. In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial subspace for a series of time windows; i.e.,  
%\methodAcronym\  
%computes the generalized coordinates
%$\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the
%minimization problems, 
\methodAcronym\ sequentially computes approximate solutions
$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$, where 
$\approxstateArgnt{n}$ is the solution to the
minimization problem

\begin{equation*}
\underset{\stateyArgnt{} \in \stspace^n}{\text{minimize } }
\mathcal{J}^n(\stateyArgnt{}), \qquad n = 1,\ldots,\nslabs,
\end{equation*}
where,
\begin{align*}
\mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
\RR{},
\end{align*}
subject to the boundary conditions. For the \spatialAcronym\ trial subspaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
\begin{align*}\
&\massArg{n}  \genstateDotArg{n}{t}  -  \basisspaceTArg{n}
\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \massArg{n} \adjointArg{n}{t} , \\
 &\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial
\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspaceArg{n}
 \adjointArg{n}{t} = \nonumber \\ 
&-\basisspaceTArg{n} \big[
\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \bigg( \mathbf{I} -
\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \bigg)
 \bigg( \basisspace \genstateDotArg{n}{t}  -
\velocity(\veloargsromn) \bigg),  &
%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
\end{align*} 
subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. The Euler--Lagrange equations expose the fact that, for \spatialAcronym\
trial subspaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
controller that enforces the residual minimization property. 
Section~\ref{sec:wls_spacetime} additionally considered \spaceTimeAcronym\ trial subspaces. For \spaceTimeAcronym\ trial subspaces, in where the generalized coordinates are no longer functions, the stationary conditions correspond to an algebraic system of equations.

>>>>

\subsection{\methodAcronym\ summary}
This section outlined the \methodAcronym\ method for model reduction. In summary, \methodAcronym\ sequentially minimizes the time continuous FOM ODE residual over the range of a space--time trial subspace for a series of time windows; i.e.,  
%\methodAcronym\  
%computes the generalized coordinates
%$\genstateArgnt{n}$, for $n = 1,\ldots,\nslabs$, as the solution to the
%minimization problems, 
\methodAcronym\ sequentially computes approximate solutions
$\approxstateArgnt{n}\in\stspace^n$, $n=1,\ldots,\nslabs$ that satisfy
\begin{equation*}
\underset{\stateyArgnt{} \in \stspace^n}{\text{minimize } }
\mathcal{J}^n(\stateyArgnt{}),
\end{equation*}
where,
\begin{align*}
\mathcal{J}^n &\vcentcolon \statey \mapsto
\frac{1}{2} \int_{\timeStartArg{n}}^{\timeEndArg{n}} \big[ \dot{\statey}(t)
- \velocity(\stateyArg{}{t},t) \big]^T \stweightingMatArg{n} \big[
\dot{\statey}(t) - \velocity(\stateyArg{}{t},t) \big] d t \\
&\vcentcolon \RR{\fomdim}\otimes \timeSpaceArg{n} \rightarrow
\RRplus,
\end{align*}
subject to the boundary conditions. For the \spatialAcronym\ trial subspaces, the stationary conditions for the residual minimization problem can be derived via the Euler--Lagrange equations and yield the system over the $n$th window for $t \in [\timeStartArg{n},\timeEndArg{n}]$,
\begin{align*}\
&\massArg{n}  \genstateDotArg{n}{t}  -  \basisspaceTArg{n}
\stweightingMatArgt{n}{t} \velocity(\veloargsromn) =  \massArg{n} \adjointArg{n}{t} , \\
 &\massArg{n} \adjointDotArg{n}{t}  + \basisspaceTArg{n} \bigg[\frac{\partial
\velocity}{\partial \statey} (\veloargsromn) \bigg]^T \stweightingMatArgt{n}{t} \basisspaceArg{n}
 \adjointArg{n}{t} = \nonumber \\ 
&-\basisspaceTArg{n} \big[
\frac{\partial \velocity}{\partial \statey}(\veloargsromn) \big]^T \stweightingMatArgt{n}{t} \big( \mathbf{I} -
\basisspaceArg{n} [\massArg{n}]^{-1} \basisspaceTArg{n} \stweightingMatArgt{n}{t} \big)
 \big( \basisspace \genstateDotArg{n}{t}  -
\velocity(\veloargsromn) \big),  &
%\genstateArg{n}{\timeStartArg{n}} = \begin{cases}
%\genstate^{n-1}(\timeEndArg{n-1}) & n=2,\ldots,\nslabs, \\
%\basisspace^T(\approxstate_0 - \stateIntercept) & n=1, \end{cases}\\ &
%\adjointArg{n}{\timeEndArg{n}} = \boldsymbol 0.
\end{align*} 
subject to the boundary conditions~\eqref{eq:lspg_bcs1}--\eqref{eq:lspg_bcs}. The Euler--Lagrange equations expose the fact that, for \spatialAcronym\
trial subspaces, \methodAcronym\ can be alternatively interpreted as a Lagrange problem from optimal control: the Galerkin method is forced by a 
controller that enforces the residual minimization property. 
Section~\ref{sec:wls_spacetime} additionally considered \spaceTimeAcronym\ trial subspaces. For \spaceTimeAcronym\ trial subspaces, in where the generalized coordinates are no longer functions, the stationary conditions correspond to an algebraic system of equations.
